---
title: "holding_place"
author: "Noah Christensen"
date: "`r Sys.Date()`"
output: html_document
---


# **HYPERION**


Hyperion data was downloaded from the **[USGS EARTH EXPLORER](https://earthexplorer.usgs.gov/)** as an *L1T Product in GeoTiff format*

Reading in L1T product

```{r,eval=FALSE}
l <- list.files(path="../../Hyperion/sevier_lake/EO1H0380332014325110PZ_1T/",
                pattern='TIF$',
                full.names=TRUE)
```


Rasterizing only working bands with the terra package

```{r,eval=FALSE}
r <- rast(l[c(8:57,77:224)])
```

## Hyperion Information
**[HYPERION BAND INFORMATION](https://developers.google.com/earth-engine/datasets/catalog/EO1_HYPERION#bands)**

```{r echo=FALSE, message=FALSE, warning=FALSE}
Hyperion_Bands <- readRDS("../../Data_Course_CHRISTENSEN/Final_Project/hyperion_band_info.rds")

# Display the DataFrame using kable with styling for a scrollable box
kbl(Hyperion_Bands) %>%
  kable_paper() %>%
  scroll_box(height = "400px")
```

## Correcting Imagery
When downloading Hyperion imagery it is stored with DN values.
Using information from the metadata you can correct this to surface reflectance.

```{r,eval=FALSE}
md <- read_lines("../../Hyperion/sevier_lake/EO1H0380332014325110PZ_1T/EO1H0380332014325110PZ_MTL_L1T.TXT")

julian_day <- as.numeric(word(md[18], 8))     #julian day
d <- (1-0.01672*cos(0.9865*(julian_day-4)))   #earth sun distance in astronomical distance
sun_elevation <- as.numeric(word(md[300],7))  #sun elevation
s <- (90-sun_elevation)                       #solar zenith angle in degrees
```

Equation to correct Hyperion DN values into surface reflectance values

```{r,eval=FALSE}
#correcting to surface reflectance
Surf_Reflectance = (pi*(r/Hyperion_Bands$Rad_Conv)*d^2)/(cos(s*pi/180)*Hyperion_Bands$Irradiance)

```


```{r include=FALSE}
Surf_Reflectance <- rast("../../Hyperion/corrected_EO1H0380332014325110PZ_1T.tif")
```

Now we can plot using three bands to create an RGB image

```{r}
Surf_Reflectance %>% 
  plotRGB(r=31,g=20,b=10,stretch = "lin")
```


## Spectral Library

Inserting spectral library from **[USGS](https://crustal.usgs.gov/speclab/QueryAll07a.php)**

Here's all the minerals in the USGS spectral library for Hyperion:
```{r echo=FALSE, message=TRUE, warning=FALSE}
minerals <- readRDS("../../Data_Course_CHRISTENSEN/Final_Project/hyperion_mineral_signatures_cleaned.rds")
# Unique values of the "mineral" column
unique_minerals <- unique(minerals$mineral)

# Set the number of columns you want
num_columns <- 7

# Determine the number of rows needed
num_rows <- ceiling(length(unique_minerals) / num_columns)

# Create a matrix to reshape the data
minerals_matrix <- matrix(unique_minerals, ncol = num_columns, byrow = TRUE)

# Convert the matrix back to a data frame
minerals_df <- as.data.frame(minerals_matrix)

# Display the table using kable
kable(minerals_df, format = "html",col.names = NULL) %>%
  kable_styling("striped", full_width = FALSE)
```


That's a lot of minerals



Google maps styling function in teams


