---
title: "Final Project"
author: "Noah Christensen"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Geologic maps serve as a valuable tool in resource exploration. 
Traditional geologic mapping is characterized by time-intensive and high-cost fieldwork, but through multispectral and hyperspectral remotely sensed images it is possible to map extensive areas instantly.




### packages used

```{r warning=FALSE, message=FALSE, echo=TRUE}
library(raster)
library(terra)
library(tidyverse)
library(janitor)
library(plotly)
library(cowplot)
library(prismatic)
library(stringr)
library(tmap)
library(mapedit)
library(rpart)
library(rpart.plot)
library(rasterVis)
library(mapview)
library(caret)
library(forcats)
```

___

# HYPERION
___


Hyperion data was downloaded from the [USGS EARTH EXPLORER](https://earthexplorer.usgs.gov/) as an *L1T Product in GeoTiff format*

Reading in L1T product

```{r,eval=FALSE}
l <- list.files(path="../../Hyperion/sevier_lake/EO1H0380332014325110PZ_1T/",
                pattern='TIF$',
                full.names=TRUE)
```


Rasterizing only working bands with the terra package

```{r,eval=FALSE}
r <- rast(l[c(8:57,77:224)])
```


*[Hyperion Band information](https://developers.google.com/earth-engine/datasets/catalog/EO1_HYPERION#bands)* collected and cleaned
```{r,eval=FALSE}
# reading in csv of hyperion band information
# removing random extra columns
# removing random extra rows
Hyperion_Bands <- read.csv("Hyperion_Bands_Wavelengths.csv") %>%
  select(-c(X,X.1,X.2,X.3))
Hyperion_Bands <- Hyperion_Bands[-(199:235),]

# removing nm and FWHM: in wavelength column
Hyperion_Bands$Wavelength <- gsub("nm","",Hyperion_Bands$Wavelength)
Hyperion_Bands$Wavelength <- gsub("FWHM:","",Hyperion_Bands$Wavelength)

# separating Wavelength and FWHM into their own columns
# separating VNIR and SWIR from band names
Hyperion_Bands <- Hyperion_Bands %>% 
  as.data.frame() %>% 
  separate(Wavelength,into = c("Wavelength_nm","FWHM_nm"),sep=",",convert = TRUE) %>%
  separate(Description,into = c("Description", "Temp", "Temp2"),sep=" ") %>%
  rename("Bands" = "Band_Name")

# removing temp columns 
Hyperion_Bands <- subset(Hyperion_Bands, select = -c(Temp, Temp2))

#adding value in new column based on Hyperion_Bands$Description
vect <- if_else(Hyperion_Bands$Description == "VNIR",40,80)

#adding vect as new column
Hyperion_Bands$Rad_Conv = vect

head(Hyperion_Bands)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
Hyperion_Bands <- readRDS("../../Data_Course_CHRISTENSEN/Final_Project/cleaned_hyperion_band_info.rds")
head(Hyperion_Bands)
```


Reading in metadata and setting up variables for Hyperion reflectance correction

```{r,eval=FALSE}
md <- read_lines("../../Hyperion/sevier_lake/EO1H0380332014325110PZ_1T/EO1H0380332014325110PZ_MTL_L1T.TXT")

julian_day <- as.numeric(word(md[18], 8))     #julian day
d <- (1-0.01672*cos(0.9865*(julian_day-4)))   #earth sun distance in astronomical distance
sun_elevation <- as.numeric(word(md[300],7))  #sun elevation
s <- (90-sun_elevation)                       #solar zenith angle in degrees
```

Equation to correct Hyperion DN values into surface reflectance values

```{r,eval=FALSE}
#correcting to surface reflectance
Surf_Reflectance = (pi*(r/Hyperion_Bands$Rad_Conv)*d^2)/(cos(s*pi/180)*Hyperion_Bands$Irradiance)
```


### Inserting spectral library from USGS

```{r echo=TRUE, warning=FALSE}
#reading in spectral mineral signatures
minerals <- list.files(path="../../usgs_spectral_library/usgs_splib07 (1)/ASCIIdata/ASCIIdata_splib07b_cvHYPERION/ChapterM_Minerals/", full.names=TRUE)

#list applying as csv and into a data frame
minerals <- minerals %>% 
  lapply(read.csv) %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(Bands = row_number())

#removing non-working bands
minerals <- minerals[c(8:57,77:224),]
```

Download Hyperion in L1T product in geotiff format


